--
title: "Survival Analysis for subtypes of HGG/DMG by splicing"
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
author: C. Savonen for ALSF CCDL, Krutika Gaonkar for D3B
date: 2019
params:
  plot_ci: TRUE
---

**Purpose:** 

Runs survival analysis models for subtypes of Diffuse astrocytic and oligodendroglial tumors with splicing types in input/HGG_AN_list.tsv. 

This module was most recently run on release-v17-20200908/pbta-histologies.tsv 

## Usage 

Uses a wrapper function (`survival_analysis`) from utils folder. 

# Setup

#### Packages and functions

Read in set up script.

```{r Set up library}
# This script contains a wrapper function that can conduct the survival analyses
source(file.path("util", "survival_models.R"))

# Magrittr pipe
`%>%` <- dplyr::`%>%`
```

#### Set up files and directories

Set up output directories. 

```{r Set up directories}
data_dir <- file.path("..", "..", "data")
results_dir <- "results"
plots_dir <- "plots"
```

Make output directories.

```{r Make output directories}
if (!dir.exists(results_dir)) {
  dir.create(results_dir)
}
if (!dir.exists(plots_dir)) {
  dir.create(plots_dir)
}
```

Declare input file paths and useful variables

```{r Set input file paths}
metadata_file <- file.path(data_dir, "pbta-histologies.tsv")
independent_primary_file <- file.path(
  data_dir,
  "independent-specimens.wgswxs.primary.tsv")

```

Declare output file paths for HGG/DMG all subtypes. 

```{r Set output file paths}
kap_meier_plot_file <- file.path(plots_dir, "survival_curve_DMG_H3_low_v_HGG_H3_high.pdf")
kap_meier_model_file <- file.path(results_dir, "survival_model_DMG_H3_low_v_HGG_H3_high.tsv")
```

## Import the metadata

```{r Read in metadata}
# choose Diffuse astrocytic and oligodendroglial tumor samples
# remove cell-lines from this analysis as well
metadata <- readr::read_tsv(
  metadata_file,
  col_types = readr::cols(
    molecular_subtype = readr::col_character())) %>%
  dplyr::filter(
      broad_histology == "Diffuse astrocytic and oligodendroglial tumor" &
      composition=="Solid Tissue") %>%
  as.data.frame() 

# read in given list of samples from Ammar
splicing_types<- readr::read_tsv("input/HGG_AN_list.tsv")

splicing_types <- splicing_types %>%
  dplyr::left_join(metadata)

```


#### Using independent_primary if multiple types per patient are found 
This way we subset the metadata to 1 single primary sample per patient ID

```{r Read in independent}

# Patients with multiple different molecular subtype might affect downstream analysis for survival so using random selection of 1-1 sample-patient matches
unique_match <- splicing_types %>%
    dplyr::group_by(Kids_First_Participant_ID) %>%
    dplyr::summarize(Kids_First_Biospecimen_ID = sample(Kids_First_Biospecimen_ID, 1)) 

# select only independent primary WGS bs_ids to get overall survival
metadata <- metadata %>%
  dplyr::filter(Kids_First_Biospecimen_ID %in%
                  unique_match$Kids_First_Biospecimen_ID) %>%
  dplyr::left_join(splicing_types[,c("sample_id","type")],suffix=c("_DNA","_RNA")) %>%
  dplyr::filter(!is.na(type)) 

```
## Kaplan-Meier for all HGG/DMG subtypes

The [Kaplan-Meier](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3059453/) 
survival curve shows the probability of survival over time.

```{r Kaplan Meier}
kap_fit <- survival_analysis(metadata,
  ind_var = "type",
  test = "kap.meier",
  metadata_sample_col = "Kids_First_Biospecimen_ID"
)
# median and 0.95CI survival days
kap_fit$model
```
Note that upper 95% bounds (0.95UCL) which are denoted "NA" are the consequence of underlying skew in the data. The upper confidence limit is therefore not defined (or may be interpreted as infinity). This can be confirmed by plotting the output. Dashed lines represent upper and lower bounds. Because the upper bound for low (blue) and high (red) never reaches as low as 0.5 along the Y-axis (Survival), its upper bound is undefined. 

```{r explain na}
plot(kap_fit$model, mark.time=FALSE, conf.int=TRUE, col=c("red", "blue"))
legend("topright", legend=c("high", "low"),col=c("red", "blue"), lty=1)

```



`survminer` package comes with a convenient plotting function which uses `ggplot2`
arguments. 
It requires the original Kaplain-Meier `survfit` object, so we will extract that with `$model`.

```{r Make survival plot}
surv_plot <- survminer::ggsurvplot(kap_fit$model,
  pval = TRUE,
  data = kap_fit$original_data,
  risk.table = TRUE,
  xlim = c(0, 3000),
  break.time.by = 500,
  ggtheme = theme_minimal(),
  risk.table.y.text.col = TRUE,
  risk.table.y.text = FALSE
)

surv_plot$plot <- surv_plot$plot +
  ggtitle(paste0("Kap_meier_Diffuse_astrocytic_and_oligodendroglial_tumor")) +
  theme(legend.position = "right")

# Make this plot a combined plot
surv_plot_all_subtype <-
  cowplot::plot_grid(surv_plot[[1]], surv_plot[[2]], nrow = 2,
                     rel_heights = c(3, 2))

# Print it out here
surv_plot_all_subtype
```

Save the plot to a file. 

```{r Save survival plot}
# We can save the plot like a normal ggplot
cowplot::save_plot(filename = kap_meier_plot_file, plot = surv_plot_all_subtype)
# Let's save the model for CI and median survial as well
readr::write_tsv(surv_median(kap_fit$model), kap_meier_model_file)
```

